Unshuffling Permutations

### 2015-06-29 article.tex
reduction
Partie algebre (mise en place)

\DeclareMathOperator{\LEFT}{\ell}
\DeclareMathOperator{\RIGHT}{r}

@@

@@

\begin{proof}[of Claim~\ref{claim:no nu_1 - nu_1 edge}]
Suppose, aiming at a contradiction that there exists
a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
\begin{tikzpicture}

\end{tikzpicture}
\qed
\end{proof}

@@

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}[
    scale=.95,
    >=stealth',
    shorten >=1pt,
    main node/.style={align=center},
    cell/.style={draw,ultra thick,fill=black!5},
    structure link/.style={line width=1.5pt},
    pattern link/.style={line width=1.5pt,black!20},
    monotone/.style={->,thick}
    ]  
    % edges
    % pi 1 - pi 2
    \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
    % pi 2 - sigma 2
    \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
    % 1
    \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
    \node (claim edge 1) at (1.5,13) {Claim\ref{}};
    % 2
    \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
    \node (claim edge 1) at (4,3.5) {Claim\ref{}};
    % 3
    \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
    \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
    % 4
    \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
    \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
    % sigma 1 - pi 1
    \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
    \node (claim edge 1) at (6.5,4) {Claim\ref{}};
    % nodes
    % 1
    \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
    \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
    \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
    \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
    % 2
    \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
    \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
    \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
    \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
    % 3
    \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
    \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
    \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
    \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
    % 4
    \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
    \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
    \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
    \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
    % sigma 1
    \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
    \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
    % pi 1
    \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
    \node [main node] (pi1) at (8.5,5.5) {$\pi'$};    
    % pi 2
    \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
    \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
    % sigma 2
    \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
    \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};  
    % grid
    \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
    % row size
    \foreach \y/\N in {0.5/N_2,1.5/N_2}
        \node at (-0.5,\y) (R\y) {$\N$};
    % }
    % column size
    \foreach \x/\N in 
    {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
        \node at (\x,-0.5) (C\x) {$\N$};
    }
  \end{tikzpicture}
\end{figure}

>>>>

\DeclareMathOperator{\LEFT}{\ell}
\DeclareMathOperator{\RIGHT}{r}
\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{std}}
\DeclareMathOperator{\SHUFFLE}{\bullet}

@@

% Définir :
% - ensemble $S$
% - standardisation $\STD$
% - notation $\pi_{|P}$ et $\pi_{|\bar P}$
% - notation $\pi^\sim$ (miroir)
% - notation $\bar \pi$ (complémentaire)
% - notation $\bar ^{-1}$ (inverse)

%%
%% ---- Algebraic issues ----
%%
\section{Algebraic issues}
\label{section:Algebraic issues}
Let use denote by $\QQ[S]$ the linear span of all permutations. We shall
define on this space a kind of shuffle product. For this purpose, we
first define a linear coproduct $\Delta$ on $\QQ[S]$ satisfying for
any permutation $\pi$,
\begin{equation} \label{equ:unshuffling_coproduct}
    \Delta(\pi) =
    \sum_{P \subseteq [|\pi|]}
    \STD\left(\pi_{|P}\right) \otimes \STD\left(\pi_{|\bar P}\right).
\end{equation}
We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
instance,
\begin{equation}
    \Delta(213) =
    \epsilon \otimes 213 + 2. 1 \otimes 12 +
    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
\end{equation}
\begin{equation}
    \Delta(1234) =
    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
    4. 123 \otimes 1 + 1234 \otimes \epsilon,
\end{equation}
\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
    \Delta(1432) & =
    \epsilon \otimes 1432 + 3. 1 \otimes 132 + 1 \otimes 321 +
    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
    321 \otimes 1 + 1432 \otimes \epsilon.
\end{split}\end{equation}
\medskip

As any coproduct, $\Delta$ leads to the definition of a product obtained
by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
for any permutation $\pi$, we have
\begin{equation}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^{(\pi)} \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^{(\pi)}$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$ denoted by
$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
permutations
$\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^{(\pi)} \; \pi.
\end{equation}
We call $\SHUFFLE$ the {\em shuffle product of permutations}. For
instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3.1432 + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
\medskip

We say that a permutation $\pi$ {\em appears} in the shuffle
$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
coefficient $\lambda_{\sigma, \nu}^{(\pi)}$ is different from zero.
Besides, a permutation $\pi$ is a {\em square} if there is a permutation
$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
Equivalently, $\pi$ is a square if in the expansion of $\Delta(\pi)$,
there is a tensor $\sigma \otimes \sigma$ with a nonzero coefficient.
\medskip

\begin{proposition}
    The shuffle product $\SHUFFLE$ of permutations is associative and
    commutative.
\end{proposition}
\begin{proof}

\end{proof}
\medskip

\begin{proposition}
    If $\pi$ is a square permutation, then $\pi^\sim$, $\bar \pi$,
    and $\pi^{-1}$ are square.
\end{proposition}
\begin{proof}

\end{proof}
\medskip

@@

\begin{proof}[of Claim~\ref{claim:no nu_1 - nu_1 edge}]
Suppose, aiming at a contradiction that there exists
a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.

\begin{tikzpicture}
  [scale=0.25]
  % nu_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu_4
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu'_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % edges
  \draw [line width=1pt]
  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
  \draw [line width=1pt]
  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
\end{tikzpicture}
\begin{tikzpicture}
  [scale=0.25]
  % nu_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu_4
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu'_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % edges
  \draw [line width=1pt]
  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
  \draw [line width=1pt]
  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
\end{tikzpicture}
\qed
\end{proof}

@@

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}[
    scale=.95,
    >=stealth',
    shorten >=1pt,
    main node/.style={align=center},
    cell/.style={draw,ultra thick,fill=black!5},
    structure link/.style={line width=1.5pt},
    pattern link/.style={line width=1.5pt,black!20},
    monotone/.style={->,thick}
    ]
    % edges
    % pi 1 - pi 2
    \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
    % pi 2 - sigma 2
    \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
    % 1
    \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
    \node (claim edge 1) at (1.5,13) {Claim\ref{}};
    % 2
    \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
    \node (claim edge 1) at (4,3.5) {Claim\ref{}};
    % 3
    \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
    \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
    % 4
    \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
    \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
    % sigma 1 - pi 1
    \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
    \node (claim edge 1) at (6.5,4) {Claim\ref{}};
    % nodes
    % 1
    \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
    \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
    \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
    \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
    % 2
    \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
    \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
    \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
    \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
    % 3
    \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
    \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
    \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
    \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
    % 4
    \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
    \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
    \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
    \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
    % sigma 1
    \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
    \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
    % pi 1
    \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
    \node [main node] (pi1) at (8.5,5.5) {$\pi'$};
    % pi 2
    \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
    \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
    % sigma 2
    \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
    \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};
    % grid
    \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
    % row size
    \foreach \y/\N in {0.5/N_2,1.5/N_2} {
        \node at (-0.5,\y) (R\y) {$\N$};
    }
    % column size
    \foreach \x/\N in
    {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
        \node at (\x,-0.5) (C\x) {$\N$};
    }
  \end{tikzpicture}
\end{figure}
